### 2025-07-19 docs/article/Sections/introduction.tex
Changes to introduction, results. Removed redundancy in 2.2.2.

\textit{Progressive Server-Side Rendering (PSSR)} combines benefits from both SSR and CSR
by streaming HTML content progressively. This technique enhances user-perceived performance by
allowing progressive rendering as data becomes available, significantly reducing the
\textit{Time to First Byte (TTFB)} and improving perceived load times compared to traditional
SSR approaches~\cite{wise2024pssr}. In this respect,
PSSR is similar to CSR in that the server initially sends a minimal HTML
document to the client and subsequently streams additional HTML fragments.
Even so, unlike CSR, PSSR retains all rendering responsibilities on the server side,
thereby reducing the load on the client. Consequently, the client does
not need to execute JavaScript or make additional requests to retrieve page content.
The streaming nature of PSSR allows users to see content progressively as it becomes
available, rather than waiting for the complete page to be rendered server-side,
thus providing a more responsive user experience with measurably lower TTFB values.

>>>>

\textit{Progressive Server-Side Rendering (PSSR)} combines benefits from both SSR and CSR
by streaming HTML content progressively. This technique enhances user-perceived performance by
allowing progressive rendering as data becomes available, significantly reducing the
\textit{Time to First Byte (TTFB)} and improving perceived load times compared to traditional
SSR approaches~\cite{wise2024pssr}. In this respect,
PSSR is similar to CSR in that the server initially sends a minimal HTML
document to the client and subsequently streams additional HTML fragments.
Even so, unlike CSR, PSSR retains all rendering responsibilities on the server side,
thereby reducing the load on the client. Consequently, the client does
not need to execute JavaScript or make additional requests to retrieve page content.
The streaming nature of PSSR allows users to see content progressively as it becomes
available, rather than waiting for the complete page to be rendered server-side,
thus providing a more responsive user experience with measurably lower TTFB values~\cite{wise2024pssr}.

### 2025-07-19 docs/article/Sections/results.tex
Changes to introduction, results. Removed redundancy in 2.2.2.

All the following benchmarks were conducted on a GitHub-hosted virtual machine
under GitHub Actions, running Ubuntu 24.04.2 LTS (Noble Numbat) with an AMD
EPYC 7763 64-Core Processor (x86\_64 architecture) operating at 3.24 GHz,
configured with 2 CPU cores and 2 threads per core, and 7.8GB of available RAM.
The system utilizes a 75GB SSD with ext4 filesystem, achieving 1.5 GB/s write
throughput in basic I/O tests. Network connectivity is provided through a 1500
MTU Ethernet interface. All tests were conducted on OpenJDK 21 (Correto build)
with the G1 garbage collector enabled by default. The following . JVM arguments
were used: \texttt{-Xms512m -Xmx16g}. Each test was executed 5 times per
configuration to ensure the reliability of the results.

@@

\subsection{Scalability results for the \texttt{Presentations} class}

@@

\subsection{Scalability results for the \texttt{Stocks} class}

@@

\subsection{Memory Consumption and Resource Utilization Analysis}

@@

\textbf{Virtual Threads vs Suspendables}

\begin{figure}[h]
     \centering
     \includegraphics[width=0.8\textwidth]{./Graphs/cpu-susp.png}
     \caption{Benchmark CPU utilization profiling with HtmlFlow-Susp in Spring WebFlux}\label{fig:cpu-susp}
\end{figure}

\begin{figure}[h]
     \centering
     \includegraphics[width=0.8\textwidth]{./Graphs/cpu-virt.png}
     \caption{Benchmark CPU utilization profiling with HtmlFlow running on Virtual Threads in Spring WebFlux}\label{fig:cpu-virt}
\end{figure}

Figures~\ref{fig:cpu-virt} and~\ref{fig:cpu-susp} show CPU utilization profiles
for the \texttt{HtmlFlow-Virtual} and \texttt{HtmlFlow-Susp} implementations,
respectively. \texttt{HtmlFlow-Virtual} reaches peak CPU usage of around 90\%
during load, with sharp transitions between idle and active phases. In
contrast, \texttt{HtmlFlow-Susp} peaks at approximately 75\% and exhibits
smoother transitions, suggesting more efficient CPU scheduling and lower
contention.

\begin{figure}[h]
     \centering
     \includegraphics[width=0.8\textwidth]{./Graphs/gc-susp.png}
     \caption{Benchmark memory utilization profiling with HtmlFlow-Susp in Spring WebFlux}\label{fig:gc-susp}
\end{figure}

\begin{figure}[h]
     \centering
     \includegraphics[width=0.8\textwidth]{./Graphs/gc-virt.png}
     \caption{Benchmark memory utilization profiling with HtmlFlow running on Virtual Threads in Spring WebFlux}\label{fig:gc-virt}
\end{figure}

The memory usage disparity is even more pronounced. As shown in
Figure~\ref{fig:gc-virt}, the Virtual Threads implementation gradually climbs
to a peak heap usage of 900MB, with large but infrequent garbage collection
events. Figure~\ref{fig:gc-susp} shows the suspendable version peaks at only
245MB, following a regular sawtooth pattern of small, frequent garbage
collections. Overall, \texttt{HtmlFlow-Virtual} allocates a total of 939MB heap
compared to just 348MB in \texttt{HtmlFlow-Susp}—a 2.7× difference.

\textbf{RxJava Blocking vs Structured Concurrency}

To further assess the efficiency of blocking versus structured concurrency, we
compared the memory behavior of RxJava's \texttt{BlockingObservableIterable}
with \texttt{PublisherAsFlow}. The blocking iterator internally relies on a
lock-based condition variable and an unbounded queue
(\texttt{SpscLinkedArrayQueue}), blocking the thread when no elements are
available. Under high concurrency, this model tends to retain large numbers of
elements in memory, especially when producers outpace consumers. This was
reflected in our profiling, where heap usage for blocking consumption showed
linear growth and irregular GC behavior.

In contrast, \texttt{PublisherAsFlow} employs coroutine-based suspension with
bounded \texttt{Channel} capacity and configurable \texttt{BufferOverflow}
strategies. It requests upstream data based on demand, leveraging structured
concurrency and cooperative cancellation.

It is important to note that our benchmark keeps data items in memory as static
objects, which represents a best-case scenario for memory allocation patterns.
In real-world applications, data typically comes from external sources such as
databases, web services, or file systems, requiring dynamic object creation and
serialization for each request. This would result in significantly higher
memory allocation rates and more frequent garbage collection cycles across all
concurrency models, with the observed differences between Virtual Threads and
suspendable coroutines likely becoming even more pronounced under such
realistic I/O-intensive workloads. While our benchmark demonstrates that
Virtual Threads can achieve comparable throughput to suspendable coroutines, or
other non-blocking paradigms, future work should further explore the impact of
dynamic data sources on memory allocation patterns and garbage collection
behavior to better understand the trade-offs between these concurrency models
in real-world scenarios.

\textbf{Summary of Trade-offs}

>>>>

This section presents the evaluation results organized into four main parts:
Section~\ref{sec:environment} details the testing environment and hardware
specifications; Section~\ref{sec:presentations-results} presents scalability
results for the \texttt{Presentations} data model;
Section~\ref{sec:stocks-results} analyzes performance with the more complex
\texttt{Stocks} data model; and Section~\ref{sec:memory-analysis} provides
detailed memory consumption and resource utilization analysis.

\subsection{Environment Specifications}
\label{sec:environment}

All benchmarks were conducted on a GitHub-hosted virtual machine under GitHub
Actions, running Ubuntu 24.04.2 LTS with an AMD EPYC 7763 64-Core Processor
(x86\_64 architecture) operating at 3.24 GHz, configured with 2 CPU cores and 2
threads per core, and 7.8 GB of available RAM. The system utilizes a 75 GB SSD
with ext4 filesystem, achieving 1.5 GB/s write throughput in basic I/O tests.
Network connectivity is provided through a 1500 MTU Ethernet interface. All
tests were conducted on OpenJDK 21 (Corretto build) with the G1 garbage
collector enabled by default. The JVM arguments used were: \texttt{-Xms512m
     -Xmx16g}. Each test was executed 5 times per configuration to ensure result
reliability.

@@

\subsection{Scalability results for the \texttt{Presentations} class}
\label{sec:presentations-results}

@@

\subsection{Scalability results for the \texttt{Stocks} class}
\label{sec:stocks-results}

@@

\subsection{Memory Consumption and Resource Utilization Analysis}
\label{sec:memory-analysis}

@@

Resource utilization data was collected using VisualVM during 40-second
benchmark runs, capturing detailed CPU and memory usage patterns under
sustained load conditions. Each profiling session monitored system performance
during the 1 to 128 concurrent user load test, providing comprehensive insights
into runtime behavior.

\subsubsection{Virtual Threads vs Structured Concurrency}

\begin{figure}[h]
     \centering
     \includegraphics[width=0.8\textwidth]{./Graphs/cpu-virt.png}
     \caption{CPU utilization profiling with HtmlFlow running on Virtual Threads in Spring WebFlux during 40-second load test}\label{fig:cpu-virt}
\end{figure}

\begin{figure}[h]
     \centering
     \includegraphics[width=0.8\textwidth]{./Graphs/cpu-susp.png}
     \caption{CPU utilization profiling with HtmlFlow-Susp in Spring WebFlux during 40-second load test}\label{fig:cpu-susp}
\end{figure}

Figures~\ref{fig:cpu-virt} and~\ref{fig:cpu-susp} show CPU utilization profiles
for the \texttt{HtmlFlow-Virtual} and \texttt{HtmlFlow-Susp} implementations,
respectively. In both cases, the initial spike in CPU usage corresponds to the
system's bootstrap and initial load phase. The subsequent gradual increase in
CPU utilization reflects the ramp-up of concurrent users during the benchmark.
\texttt{HtmlFlow-Virtual} exhibits a distinctive bell-shaped curve, rapidly
climbing from near-zero to peak CPU usage of around 90\% during the initial
load phase, maintaining high utilization during sustained load, and then
dropping sharply back to baseline. The profile shows more aggressive CPU
consumption with steep ramp-up and ramp-down phases. In contrast,
\texttt{HtmlFlow-Susp} displays a more gradual and controlled resource
utilization pattern, with a smoother ascent to a lower peak of approximately
75\%, followed by a more gentle descent. This smoother profile suggests more
efficient CPU scheduling and better resource management under load. The GC
activity indicators in both profiles remain relatively low throughout the test
duration, confirming that garbage collection overhead does not significantly
impact CPU utilization in either approach.

\begin{figure}[h]
     \centering
     \includegraphics[width=0.8\textwidth]{./Graphs/gc-virt.png}
     \caption{Memory utilization and garbage collection behavior with HtmlFlow running on Virtual Threads in Spring WebFlux during 40-second load test}\label{fig:gc-virt}
\end{figure}

\begin{figure}[h]
     \centering
     \includegraphics[width=0.8\textwidth]{./Graphs/gc-susp.png}
     \caption{Memory utilization and garbage collection behavior with HtmlFlow-Susp in Spring WebFlux during 40-second load test}\label{fig:gc-susp}
\end{figure}

The memory usage disparity is more pronounced. Memory utilization data was
collected using VisualVM profiling during the same 40-second benchmark runs. As
shown in Figure~\ref{fig:gc-virt}, the Virtual Threads implementation gradually
climbs to a peak heap usage of 900MB, with large but infrequent garbage
collection events. Figure~\ref{fig:gc-susp} shows the suspendable version peaks
at only 245MB, following a regular sawtooth pattern of small, frequent garbage
collections. Overall, \texttt{HtmlFlow-Virtual} allocates a total of 939MB heap
compared to just 348MB in \texttt{HtmlFlow-Susp}—a 2.7× difference.

The higher observed heap usage with Virtual Threads stems from how they manage
concurrency. Virtual Threads allocate a full stack per thread, and when they
are suspended (e.g., during blocking I/O), their entire call stack is unmounted
and copied to the heap, allowing the physical carrier thread to be reused.
Kotlin coroutines, by contrast, also store their execution state on the heap
when suspended — but the compiler transforms coroutine code into a state
machine, so only the minimal necessary state is saved rather than a full stack.
This results in significantly lower memory overhead per suspended task.

It is important to note that our benchmark keeps data items in memory as static
objects, which represents a best-case scenario for memory allocation patterns.
In real-world applications, data typically comes from external sources such as
databases, web services, or file systems, requiring dynamic object creation and
serialization for each request. This would result in significantly higher
memory allocation rates and more frequent garbage collection cycles across all
concurrency models, with the observed differences between Virtual Threads and
suspendable coroutines likely becoming even more pronounced under such
realistic I/O-intensive workloads.

\subsubsection{Summary of Trade-offs}
